E-graphs With Bindings

### 2025-01-08 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
edit on soundness and completeness and draft for closed

%!LW recipe=pdflatex -> bibtex -> pdflatex * 2
\documentclass[acmsmall, screen, nonacm]{acmart}

\newcommand*{\ONECOLUMN}{}

% \input{../preamble}

\usepackage{hyperref}

@@

\usepackage{import}

\newcommand\HypI[1]{\textbf{HypI}(#1)}
\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
% conflict
\newcommand{\hashtag}{{\#}}
\newcommand{\consistency}{{\smile}}
% There is already a remark in the preamble, but it does
% not work for some reason
\newcommand\obj[1]{{\textbf{obj}(#1)}}

\usepackage{xifthen}
\ifthenelse{\isundefined{\ismain}}{
  \newcommand{\ismain}{0}
}{}

\usepackage{enumitem}

\newcommand{\ignora}[1]{ }

% \newcounter{thm}

% % \ignora{
% \usepackage{aliascnt}
% % alias-counters
% \newaliascnt{definition}{thm}
% \newaliascnt{proposition}{thm} 
% \newaliascnt{lemma}{thm}
% \newaliascnt{corollary}{thm}
% \newaliascnt{conjecture}{thm}
% \newaliascnt{remark}{thm}
% \newaliascnt{example}{thm}
% \newaliascnt{examples}{thm}
% \newaliascnt{assumption}{thm}
% \newaliascnt{construction}{thm}
% \newaliascnt{claim}{thm}

% % \theoremstyle{plain}
% \theoremstyle{definition}

% \newtheorem{theorem}[thm]{Theorem}
% \newtheorem{proposition}[proposition]{Proposition} 
% \aliascntresetthe{proposition}
% \newtheorem{lemma}[lemma]{Lemma}
% \aliascntresetthe{lemma}
% \newtheorem{corollary}[corollary]{Corollary}
% \aliascntresetthe{corollary}
% \newtheorem{conjecture}[conjecture]{Conjecture}
% \aliascntresetthe{conjecture}
% \newtheorem{remark}[remark]{Remark}
% \aliascntresetthe{remark}
% \newtheorem{assumption}[assumption]{Assumption}
% \aliascntresetthe{assumption}
% \newtheorem{construction}[construction]{Construction}
% \aliascntresetthe{construction}
% \newtheorem{claim}[claim]{Claim}
% \aliascntresetthe{claim}

% \theoremstyle{definition}
% \newtheorem{definition}[definition]{Definition}
% \aliascntresetthe{definition}
% \newtheorem{example}[example]{Example}
% \aliascntresetthe{example}
% \newtheorem{examples}[examples]{Examples}
% \aliascntresetthe{examples}
% }

@@

\usepackage{cleveref} % after hyperref!

\input{../macros}
\input{../sample.tikzstyles}
\input{../hypergraph.tikzstyles}
\input{../hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

\newcommand\enriched[1]{{\overline{\mathcal{#1}}}}

\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
% \setcopyright{acmlicensed}
% \copyrightyear{2024}
% \acmYear{2024}
% \acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.

%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
% \acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\begin{abstract}
  These notes contain the foundations of categories enriched over a category of semilattices and the respective string diagrams for such categories.
  It introduces the required notions and carefully checks that all properties are satisfied.
\end{abstract}

\begin{document}

\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}

@@

\begin{section}{Enriched Categories}

\begin{definition}[Definition 6.2.1~\cite{Borceux_1994}]
  Let $\mathcal{V} = (\mathcal{V}_{0}, I_{\mathcal{V}}, \otimes, l, r, a)$ be a monoidal category where $l$ and $r$ are left and right unitors.
  We define a $\mathcal{V}$-enriched category $\enriched{C}$ consisting of the following data
  \begin{enumerate}
    \item a class of objects $\obj{\enriched{C}}$
    \item for every pair of objects $A,B \in \obj{\enriched{C}}$ --- an (hom-)object $\obj{\enriched{C}}(A,B) \in \mathcal{V}$
    \item for every triple of objects $A,B,C \in \obj{\enriched{C}}$ --- a composition morphism in $\mathcal{V}$
    \[
    c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)
    \]
    \item for every object $A \in \obj{\enriched{C}}$ --- a unit morphism
    \[
      u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)
    \]
  \end{enumerate}
  such that the following diagrams commute (associativity and unitality)
  \[
  \begin{tikzcd}
    {(\enriched{C}(A,B) \otimes \enriched{C}(B,C)) \otimes \enriched{C}(C,D)} \arrow[rr, "{c_{A,B,C} \otimes 1}"] \arrow[d, "{a_{\enriched{C}(A,B),\enriched{C}(B,C),\enriched{C}(C,D)}}"']   &  & {\enriched{C}(A,C) \otimes \enriched{C}(C,D)} \arrow[dd, "{c_{A,C,D}}"]                                                                                     \\
    {\enriched{C}(A,B) \otimes (\enriched{C}(B,C) \otimes \enriched{C}(C,D))} \arrow[d, "{1\otimes c_{B,C,D}}"']                                             &  &                                                                                                                                       \\
    {\enriched{C}(A,B) \otimes \enriched{C}(B,D)} \arrow[rr, "{c_{A,B,D}}"]                                                                       &  & {\enriched{C}(A,D)}                                                                                                                             
  \end{tikzcd}
  \]

  \[
  \begin{tikzcd}
    {I_{\mathcal{V}} \otimes \enriched{C}(A,B)} \arrow[dd, "u_{A} \otimes 1"'] \arrow[r, "{l_{\enriched{C}(A,B)}}"]                                             & {\enriched{C}(A,B)} \arrow[dd, "{1_{\enriched{C}(A,B)}}"] & {\enriched{C}(A,B) \otimes I_{\mathcal{V}}} \arrow[l, "{r_{\enriched{C}(A,B)}}"'] \arrow[dd, "1 \otimes u_{B}"]                                                           \\
    &                                     &                                                                                                                                       \\
{\enriched{C}(A,A) \otimes \enriched{C}(A,B)} \arrow[r, "{c_{A,A,B}}"]                                                                        & {\enriched{C}(A,B)}                            & {\enriched{C}(A,B) \otimes \enriched{C}(B,B)} \arrow[l, "{c_{A,B,B}}"']                                                                                    
\end{tikzcd}
  \]
\end{definition}

\begin{remark}
  We will occasionally use $c_{A,B,C}$ as $;$, \textit{i.e.}, instead of $c_{A,B,C}(f,g)$ we will write $f ; g$ when it makes sense to talk about components of hom-objects.
\end{remark}

\begin{remark}

Requiring that a category $\mathcal{V}$ is monoidal is the most basic requirement for enriched category theory.
To make the notion of enriched category more useful, in particular to be able to talk about a free enrichment, we will further require that $\mathcal{V}$ is a closed symmetric monoidal category.
\end{remark}

\begin{example}
When $\mathcal{V} = \catname{Set}$ we end up with the usual definition of a category.
$c_{A,B,C}$ becomes the usual morphism composition ($f;g$ rather than $g \circ f$) and $u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)$ picks out the identity morphism from the homset $\enriched{C}(A,A)$ by matching the only element of a singleton set $I \cong \{*\} \mapsto 1_{A}$ and component-wise the diagrams above commute.
\end{example}

TODO: another example

\begin{definition}[Semilattice]
  A \textit{semilattice} is a set equipped with an operation that we denote as $+$ which is associative, commutative and idempotent.
\end{definition}

Note that we do not require the existence of a unit for $+$. 
Semilattices that satisfy this extra requirement are sometimes called \textit{bounded}, i.e., they are idempotent commutative monoids.

\begin{definition}[Semilattice homomorphism]

A homomorphism between two semilattices $S_{1}$ and $S_{2}$ is a map $h$ that respects $+$.
That is, for all $s,s' \in S_{1}$, $h(s +_{S_{1}} s') = h(s) +_{S_{2}} h(s')$.
\end{definition}

\begin{definition}[Category of semilattices]

Semilattices with their respective homomorphisms form a category that we denote $\catname{SLat}$.
\end{definition}

\begin{proposition}
  $\catname{SLat}$ is a closed symmetric monoidal category.
\end{proposition}
\begin{proof}
  The tensor product of two semilattices $S_{1}$ and $S_{2}$ is defined as follows.
  $S_{1} \otimes S_{2}$ consists of pairs $(s_1,s_2)$ $s_{1} \in S_{1}$, $s_{2} \in S_{2}$ quotiented by commutativity, idempotence and associativity and additionally by the following relations
  \begin{itemize}
    \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
    \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
  \end{itemize}

  The unit for this tensor product is $I = \{*\}$ --- a one-element semilattice.
  Clearly $S \otimes I \cong S$ by mapping $(s,*) \mapsto s$ for all $s \in S$ and vice versa.
  The symmetry, associators and unitors are then obvious morphisms.
  Finally, the category is closed since the set of homomorpisms between two semilattices is a semilattice by defining $(f + g)(x)$ as $f(x) + g(x)$.
  $f + g$ is a homomorphism since $(f + g)(x+y) = f(x+y) + g(x+y) = f(x) + f(y) + g(x) + g(y) = f(x) + g(x) + f(y) + g(y) = f(x+y) + g(x+y)$.
\end{proof}

This makes $\catname{SLat}$ a suitable base for enrichment and allows us to take $\mathcal{V} = \catname{SLat}$.
Before we proceed we need to define what an enriched functor is and define what it means to freely enrich a category over $\catname{SLat}$.

\begin{definition}[Isomorphism in $\enriched{C}$. Lemma 3.5.12~\cite{Riehl_2014}]

Let $\enriched{C}$ be a $\mathcal{V}$-enriched category.
Given two objects $A,B \in \enriched{C}$ and a morphism $f : I_{\mathcal{V}} \to \enriched{C}(A,B)$, we call $f$ an isomorphism if there is a $g : I_{\mathcal{V}} \to \enriched{C}(B,A)$ such that
\[
  I_{\mathcal{V}} \cong I_{\mathcal{V}} \otimes I_{\mathcal{V}} \xrightarrow{f \otimes g} \enriched{C}(A,B) \otimes \enriched{C}(B,A) \xrightarrow{c_{A,B,A}} \enriched{C}(A,A)
\]
equals $u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)$.
\end{definition}

% TODO: I took this definition from \url{https://mathoverflow.net/questions/326575/isomorphisms-in-enriched-categories#:~:text=equals%20id%3A1%E2%86%92,C(X%2CY).} 
% Need a proper reference. Riehl, for example, only defines isomorphism of $A$ $B$ in terms of objects of the underlying category.

\begin{definition}
  Let $\enriched{C}$ and $\enriched{D}$ be two $\mathcal{V}$-categories.
  A $\mathcal{V}$-enriched functor $F : \enriched{C} \to \enriched{D}$ is defined by the following data.
  \begin{enumerate}
    \item A mapping $F : \obj{\enriched{C}} \to \obj{\enriched{D}}$
    \item An object-indexed family of morphisms in $\mathcal{V}$ $F_{A,B} : \enriched{C}(A,B) \to \enriched{D}(FA,FB)$
  \end{enumerate}
  such that the following diagrams commute

  \[
  \begin{tikzcd}
	{\enriched{C}(A,A') \otimes \enriched{C}(A',A'')} && {\enriched{C}(A,A'')} \\
	\\
	{\enriched{D}(FA,FA') \otimes \enriched{D}(FA',FA'')} && {\enriched{D}(FA,FA'')}
	\arrow["{c_{A,A',A''}}"', from=1-1, to=1-3]
	\arrow["{F_{A,A'} \otimes F_{A',A''}}", from=1-1, to=3-1]
	\arrow["{F_{A,A''}}"', from=1-3, to=3-3]
	\arrow["{c_{FA,FA',FA''}}"', from=3-1, to=3-3]
\end{tikzcd}
\]

\[
\begin{tikzcd}
	I_{\mathcal{V}} && {\enriched{C}(A,A)} \\
	\\
	&& {\enriched{D}(FA,FA)}
	\arrow["{u_{A}}", from=1-1, to=1-3]
	\arrow["{u_{FA}}"', from=1-1, to=3-3]
	\arrow["{F_{A,A}}", from=1-3, to=3-3]
\end{tikzcd}
\]
\end{definition}

\begin{definition}
  Given two $\mathcal{V}$-functors $F,G : \enriched{C} \to \enriched{D}$ a $\mathcal{V}$-natural transformation is family of morphisms indexed by objects in $\enriched{C}$, $\alpha_{A} : I_{\mathcal{V}} \to \enriched{D}(FA,GA)$ in $\mathcal{V}$ such that the following diagram commutes
  \[
  \begin{tikzcd}
	& {\enriched{C}(A,A')\otimes I_{\mathcal{V}}} && {\enriched{D}(FA,FA') \otimes \enriched{D}(FA',GA')} \\
	{\enriched{C}(A,A')} &&&& {\enriched{D}(FA,GA')} \\
	& {I_{\mathcal{V}} \otimes \enriched{C}(A,A')} && {\enriched{D}(FA,GA) \otimes \enriched{D}(GA,GA')}
	\arrow["{G_{A,A'}\otimes \alpha_{A'}}"', from=1-2, to=1-4]
	\arrow["{c_{FA,FA',GA'}}"', from=1-4, to=2-5]
	\arrow["{r^{-1}}", from=2-1, to=1-2]
	\arrow["{l^{-1}}", from=2-1, to=3-2]
	\arrow["{\alpha_{A} \otimes F_{A,A'}}", from=3-2, to=3-4]
	\arrow["{c_{FA,GA,GA'}}", from=3-4, to=2-5]
\end{tikzcd}
\]
\end{definition}

\begin{remark}
  When $\mathcal{V}$ is $\catname{Set}$, a family of morphisms $\alpha_{A} : I_{\mathcal{V}} \to \enriched{C}(FA,GA)$ is defined by mapping $* \mapsto \alpha'_{A}$ where $\alpha'_{A} : FA \to GA$ is usual natural transformation.
  We can then trace the diagram to get

\[\begin{tikzcd}
	& {f \otimes *} && {Gf \otimes \alpha'_{A'}} \\
	f &&&& {Gf;\alpha'_{A} = \alpha'_{A};Ff} \\
	& {* \otimes f} && {\alpha'_{A} \otimes Ff}
	\arrow[from=1-2, to=1-4]
	\arrow[from=1-4, to=2-5]
	\arrow[from=2-1, to=1-2]
	\arrow[from=2-1, to=3-2]
	\arrow[from=3-2, to=3-4]
	\arrow[from=3-4, to=2-5]
\end{tikzcd}\]

which is the usual naturality condition.

\end{remark}

\begin{definition}
Forgetful functor $U : \catname{SLat} \to \catname{Set}$ is given by $\catname{SLat}(I_{\catname{SLat}, -}) : \catname{SLat} \to \catname{Set}$.
\end{definition}

Intuitively, the above functor returns the underlying set of a given semilattice $S$ as each morphism from $\{*\} \to S$ picks out an element of $S$.

\begin{proposition}
  The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
\end{proposition}
\begin{proof}
  The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
  The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
  The adjunction is then given by the following natural isomorphisms
  \[
  \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A, U(B))
  \]
  The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
  Furthermore, we have, 
  \[
  F(I) \cong I_{\catname{SLat}}
  \]
  and 
  \begin{align*}
  F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
        &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
        &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
        &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
        &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
        &\cong F(A \times B)
  \end{align*}
In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
\end{proof}

\begin{remark}
  The above proposition in fact works for any symmetric monoidal closed category $\mathcal{V}$ picked as a base for enrichment (Proposition 6.4.6~\cite{Borceux_1994}) 
\end{remark}

\begin{remark}

  Free semilattice on a set $S$ can be though of as a set of non-empty subsets of $S$.
\end{remark}

\begin{definition}[Proposition 6.4.7~\cite{Borceux_1994}]
\label{def:free_slat}
  Let $\mathcal{C}$ be a small category.
  Freely enriched $\catname{SLat}$-category $\enriched{C}$ is defined as follows
  \begin{enumerate}
    \item $\obj{\enriched{C}} = \obj{\mathcal{C}}$;
    \item $\enriched{C}(A,B) = F(\mathcal{C}(A,B))$, where $F$ is the left adjoint to $U : \catname{SLat} \to \catname{Set}$;
    \item $c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)$ is the composite $\cong ; F(;)$ below
    \[
    F(\mathcal{C}(A,B)) \otimes F(\mathcal{C}(B,C)) \xrightarrow{\cong} F(\mathcal{C}(A,B) \otimes \mathcal{C}(B,C)) \xrightarrow{F(;)} F(\mathcal{C}(A,C))
    \] 
    where the rightmost arrow is the image of the ordinary composition;
    \item $u_{A} : I_{\catname{SLat}} \to \enriched{C}(A,A)$ is the composite
    \[
    I_{\catname{SLat}} \xrightarrow{\cong} F(I) \to F(\mathcal{C}(A,A))
    \]
    where the rightmost arrow is $* \mapsto 1_{A}$.
  \end{enumerate}
\end{definition}

This is again a specialisation of a more general definition of a free $\mathcal{V}$-category (Proposition 6.4.7~\cite{Borceux_1994}).

We will then consider \textit{only} such freely-enriched categories.

\begin{remark}
  Suppose $\mathcal{C}(A,B) = \{f,g\}$ and $\mathcal{C}(B,C) = \{h\}$, then $f + g \in \enriched{C}(A,B)$ and in particular $(f+g);h = f;h + g;h$ since $c_{A,B,C}$ is a semilattice homomorphism.
\end{remark}

\begin{proposition}
\label{prop:functoriality_preserved}
  A functor $G : \mathcal{C} \to \mathcal{D}$ induces an enriched functor $\overline{G} : \enriched{C} \to \enriched{D}$.
\end{proposition}
\begin{proof}
  We define
  \begin{itemize}
    \item $\overline{G}(A) = G(A)$ for all $A \in \enriched{C}$
    \item $\overline{G}_{A,B}(\enriched{C}(A,B)) = F(G_{A,B}(\mathcal{C}(A,B)))$
  \end{itemize}
  The functoriality follows from the following diagrams
\[
\adjustbox{width=\linewidth}{
\begin{tikzcd}
	{\enriched{C}(A,A')\otimes \enriched{C}(A',A'')} && {\enriched{C}(A,A'')} && {(f, g)} && {f;g} \\
	\\
	{\enriched{D}(GA,GA') \otimes \enriched{D}(GA',GA'')} && {\enriched{D}(GA,GA'')} && {(Gf,Gg)} && {Gf;Gg = G(f;g)}
	\arrow[from=1-1, to=1-3]
	\arrow[from=1-1, to=3-1]
	\arrow[from=1-3, to=3-3]
	\arrow[from=1-5, to=1-7]
	\arrow[from=1-5, to=3-5]
	\arrow[from=1-7, to=3-7]
	\arrow[from=3-1, to=3-3]
	\arrow[from=3-5, to=3-7]
\end{tikzcd}}
\]
where the right-hand diagram is a component render of the left-hand diagram.
It commutes because $G$ is a functor. 
By freeness, the commutativity of more complex components follows from the commutativity for generators.
The following diagrams commute for the same reason.
\[\begin{tikzcd}
	I_{\catname{SLat}} && {\enriched{C}(A,A)} && {*} && {1_{A}} \\
	\\
	&& {\enriched{D}(GA,GA)} &&&& {G(1_{A}) = 1_{G(A)}}
	\arrow[from=1-1, to=1-3]
	\arrow[from=1-1, to=3-3]
	\arrow[from=1-3, to=3-3]
	\arrow[from=1-5, to=1-7]
	\arrow[from=1-5, to=3-7]
	\arrow[from=1-7, to=3-7]
\end{tikzcd}\]
\end{proof}

\begin{remark}
  Again, because $\overline{G}_{A,B}$ is a semilattice homomorphism, we have $\overline{G}(f+g) = \overline{G}(f) + \overline{G}(g)$ for any $\catname{SLat}$-enriched functor $\overline{G}$.
\end{remark}

\begin{example}
  Let $\mathcal{C} = (\mathcal{C}_{0},I,\otimes, l, r, \alpha)$ be a monoidal category and consider the bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to \mathcal{C}$ on $\mathcal{C}$.
  It induces an enriched monoidal category $\enriched{C} = (\enriched{C}_{0}, I, \enriched{\otimes}, \overline{l}, \overline{r}, \enriched{\alpha})$ such that
  \begin{enumerate}
    \item $A \overline{\otimes} B \coloneq A \otimes B$;
    \item $\overline{l} : A \overline{\otimes} I \to A \coloneq I \to \enriched{C}(A \overline{\otimes} I, I)$, i.e. $\overline{l}$ picks out $l$ out of $\enriched{C}(A \overline{\otimes} I, I)$;
    \item similarly for $\overline{r}$, $\overline{\alpha}$.
  \end{enumerate}
  $\otimes_{A,B,C,D} : \mathcal{C}(A,B) \times \mathcal{C}(C,D) \to \mathcal{C}(A \otimes C, B \otimes D)$ becomes $\overline{\otimes}_{A,B,C,D} : \enriched{C}(A,B) \times \enriched{C}(C,D) \to \enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D)$.
  In particular, $f + g \;\overline{\otimes}\; h + k = (f \;\overline{\otimes}\; h + k) + (g \;\overline{\otimes}\; h + k) = (f \;\overline{\otimes}\; k) + (f \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; k)$, where $+$ binds tighter than $\overline{\otimes}$.
  \question{I need to show that $\overline{\otimes}$ is a bifunctor. Functoriality in each argument does not entail bifunctoriality. This is the difference between monoidal and pre-monoidal categories.
  }
\end{example}

>>>>

\documentclass[peerreviewcls]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%Template version as of 6/27/2024

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

@@

@@

% some macros that maybe should be moved elsewhere

\newcommand\HypI[1]{\textbf{HypI}(#1)}
\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
% conflict
\newcommand{\hashtag}{{\#}}
\newcommand{\consistency}{{\smile}}

\input{../../macros}
\input{../../sample.tikzstyles}
\input{../../hypergraph.tikzstyles}
\input{../../hypergraph.tikzdefs}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{E-graphs With Bindings
}

% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% }

\author{Anonymous}

@@

\begin{abstract}
This is abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert.
\end{IEEEkeywords}

\section{Introduction}
E-graphs~\cite{EggPaper} are a data structure designed to facilitate equality reasoning in the context of term rewriting systems, \textit{e.g.}, for program optimisation.
They aim to solve the phase-ordering problem that arises when performing term rewriting, that is, the problem of one rewrite blocking the application of another rewrite due to removing the redex from the initial term.
They do this by making the process of applying rewrite rules non-destructive: an application of a rewrite rule $\langle l, r \rangle$ simply adds the right-hand side $r$ to the e-graph that has a term with a redex for $l$ such that the new e-graph contains the original term and the $r$ at the same time in a compact manner by utilising sharing as much as possible.
This way the application of another rule $\langle l, r' \rangle$ would not be blocked as the residual e-graph still contains the redex with $l$ which would have been blocked should we simply replaced all the occurrences of $l$ in the initial term with $r$.
E-graphs have been around for a while since being introduced in the 80s~\cite{nelson1980techniques} and yet there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.

There are domains, however, where the effective usage of e-graphs requires \textit{ad hoc} tweaks to achieve the desired performance, namely domains that include variable \textit{binding} and the closely related notion of \textit{substitution}, for example, lambda calculus with $\beta$-reduction.
The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{de Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a substitution of equivalence classes that represent a lot of terms at once.
This work shows how these two problems can be uniformly solved with \textit{string diagrams} by building on top of the recent result of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} that provided a categorical semantics for e-graphs by showing a correspondence between e-graphs and string diagrams for Cartesian semilattice-enriched symmetric monoidal categories or concrete combinatorial structures called \textit{e-hypergraphs} that encode them.
It proceeds in doing so by showing how e-graphs with bindings can be realized as string diagrams (or e-hypergraphs) for closed (Cartesian) semilattice-enriched symmetric monoidal categories.

\subsection{E-graphs}

E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes.
The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.

\begin{figure*}
\[
\adjustbox{scale=0.7}{
\tikzfig{figures/e-graph-example}
}
\]
\caption{E-graph example (top) and its equivalent string diagram representation (bottom)}
\label{fig:e-graph-example}
\end{figure*}

Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph as developed in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}. 
The initial term, $(a*2)/2$ (which corresponds to subfigure (a)), is already represented efficiently in the e-graph by sharing the node 2.
The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node
\scalebox{0.3}{
	\begin{tikzpicture}
		\begin{pgfonlayer}{nodelayer}
			\node [style=vertex] (0) at (1.5, 3.5) {};
			\node [style=none] (1) at (1.5, 3) {};
			\node [style=none] (2) at (1, 4) {};
			\node [style=none] (3) at (2, 4) {};
		\end{pgfonlayer}
		\begin{pgfonlayer}{edgelayer}
			\draw (0) to (1.center);
			\draw [bend right=45] (0) to (3.center);
			\draw [bend left=45] (0) to (2.center);
		\end{pgfonlayer}
	\end{tikzpicture}
}
.

Nodes of an e-graph are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs rooted at these boxes.
Initially each equivalence class has a single node (subgraph).
The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary (containing more than one node or subgraph) equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
Explicit discarding (weakening) nodes, depicted with black dots with a single wire inside the dashed boxes 
\scalebox{0.4}{
	\begin{tikzpicture}
		\begin{pgfonlayer}{nodelayer}
			\node [style=vertex] (0) at (1.5, 3.5) {};
			\node [style=none] (1) at (1.5, 3) {};
		\end{pgfonlayer}
		\begin{pgfonlayer}{edgelayer}
			\draw (0) to (1.center);
		\end{pgfonlayer}
	\end{tikzpicture}
}, indicate which of the class-level edges are connected to which nodes inside the class.

The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules.

\textcolor{red}{Add example showing that using De Brujin indices with e-graphs breaks sharing}

\textcolor{red}{Add example showing $\beta$-reduction in e-graphs compared to string diagrams}

\begin{figure*}
    \[
        % \hspace{1.3cm}
        \adjustbox{width=\textwidth}{
        \tikzfig{../../figures/categorical-semantics/e-graph-substitution}
        }
    \]
    \caption{E-graph explicit substitution example.}
    \label{fig:e-graph-substitution}
  \end{figure*}

  \begin{figure*}
    \[
        % \hspace{1.3cm}
        \adjustbox{width=\textwidth}{
        \tikzfig{../../figures/categorical-semantics/e-string-substitution-2}
        }
    \]
    \caption{E-string diagrammatic substitution example.}
    \label{fig:e-string-substitution}
  \end{figure*}

\section{Categorical semantics}

Free semilattice enrichment of a closed SMC.

\section{Combinatorial semantics}
